Measurement Bias from Address Aliasing

### 2014-11-19 paper.tex
Split motivation into separate figures

%\pgfplotstableread{bin/microkernel-cycles-core2.dat}{\microkernelcoretwotable}
\pgfplotstableread{bin/microkernel-core.dat}{\tblmicrokernelcore}
\begin{figure}[t]
\begin{minipage}[t]{0.40\textwidth}
  \vspace{3mm}
  \begin{lstlisting}[frame=tb]
  static int i, j, k;
  int main() {
      int g = 0, inc = 1;
      for (; g < 65536; g++) {
          i += inc;
          j += inc;
          k += inc;
      }
      return 0;
  }
  \end{lstlisting}
\end{minipage}
\begin{minipage}[t]{0.55\textwidth}
  \begin{flushleft}
    \begin{tikzpicture}
      \begin{axis}[
          title = Core,
          width = 7.8cm,
          height = 4.5cm,
          font = \footnotesize,
          xlabel=Bytes added to environment,
          ylabel=Cycles,
          domain = 0:4096,
          xtick = {0,1024,...,4096},
          xmin = 0,
          xmax = 4096,
          cycle list name = exotic
        ]
        \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \tblmicrokernelcore ;
      \end{axis}
    \end{tikzpicture}
  \end{flushleft}
\end{minipage}
\label{fig:motivation}
\caption{Microkernel first presented by \citeN{Mytkowicz:2009:WrongData}, showing bias to environment size. Notice the clear spike at around 3200 bytes, where the number of cycles executed nearly doubles. Measuring performance in this configuration would give a biased result.}
\end{figure}

@@

\subsection{Microkernel Analysis}
\label{sec:microkernel}
To illustrate how address aliasing can cause bias, we revisit the example first presented in ``Producing Wrong Data Without Doing Anything Obviously Wrong!'' by \citeN{Mytkowicz:2009:WrongData}, a small C program reproduced here in Figure~\ref{fig:motivation}.
This example is interesting for several reasons; the bias effects are significant and easily reproducible, while the code itself is simple and straightforward to analyze.
Still, no satisfactory explanation as to what actually causes bias was given in the original paper.
The following outlines our data and analysis of this program conducted on the ``Haswell'' configuration.
Similar experiments were conducted on the other experimental setups listed in Table~\ref{tab:machines} as well, all arriving at the same conclusion.

>>>>

\begin{figure}[t]
  \begin{lstlisting}[frame=single, xleftmargin=.01\textwidth, xrightmargin=.01\textwidth]
static int i, j, k;
int main() {
    int g = 0, inc = 1;
    for (; g < 65536; g++) {
        i += inc;
        j += inc;
        k += inc;
    }
    return 0;
}
  \end{lstlisting}
  \caption{\label{fig:microkernel}Microkernel first presented by \citeN{Mytkowicz:2009:WrongData}, showing bias to environment size.}
\end{figure}

%\pgfplotstableread{bin/microkernel-cycles-core2.dat}{\microkernelcoretwotable}
\pgfplotstableread{bin/microkernel-core.dat}{\tblmicrokernelcore}
\begin{figure}[t]
  \begin{tikzpicture}
    \begin{axis}[
        title = Core,
        width = 7.8cm,
        height = 4.5cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:4096,
        xtick = {0,1024,...,4096},
        xmin = 0,
        xmax = 4096,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \tblmicrokernelcore ;
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:motivation}Performance counter data running the micro-kernel from Figure~\ref{fig:microkernel} under different environment sizes. Notice the clear spike at around 3200 bytes, where the number of cycles executed nearly doubles. Measuring performance in this configuration would give a biased result.}
\end{figure}

@@

\subsection{Microkernel Analysis}
\label{sec:microkernel}
To illustrate how address aliasing can cause bias, we revisit the example first presented in ``Producing Wrong Data Without Doing Anything Obviously Wrong!'' by \citeN{Mytkowicz:2009:WrongData}, a small C program reproduced here in Figure~\ref{fig:microkernel}.
This example is interesting for several reasons; the bias effects are significant and easily reproducible, while the code itself is simple and straightforward to analyze.
Still, no satisfactory explanation as to what actually causes bias was given in the original paper.
The following outlines our data and analysis of this program conducted on the ``Haswell'' configuration.
Similar experiments were conducted on the other experimental setups listed in Table~\ref{tab:machines} as well, all arriving at the same conclusion.
